Autocorrelation in Sparse Distributed Memory

### 2018-08-22 Chapters/hathor-bitcoin-math.tex
Merge branch 'master' of https://github.com/Alex-Linhares/msbrogli-phd

The primary objective of this chapter is to increase the understanding of the Bitcoin through mathematical tools.

@@

In the Bitcoin's protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average. Suppose that the Bitcoin's network is able to calculate $H$ hashes per second --- $H$ is the total hash rate of the network. The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block. So, the rule of finding a new block every 10 minutes ($\eta = 600$ seconds) --- on average --- leads to the following equation: $\mathbf{E}(T) = \eta = 600$. So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = \eta = 600 \Rightarrow p = \frac{1}{\eta H}$. Finally, $\mathbf{E}(X) = \eta H$, $\mathbf{E}(T) = \eta$, $\mathbf{V}(X) = (\eta H)^2 - \eta H$, and $\mathbf{V}(T) = \eta^2 - \eta/H$.

The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\eta H} \right)^{tH}$. But, as the Bitcoin's network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\eta}$.

@@

Bitcoin peer-to-peer network is a gossip network, where miners are semi-randomly connected to each other, and each miner sends all information it receives to all its peers. According to \citet{decker2013information}, the average time for a new block propagate over the network is 12.6 seconds, while the 95\% percentile is 40 seconds, which indicates a long-tail distribution. \citet{bitcoinstats} has measured the propagation time between 2013 and 2017. During 2017, the worst daily 90\% percentile was 21 seconds. Notice that both results may not be contradictory because Bitcoin network is continuously evolving.

@@

@@

Let's say that the network's total hash rate has suddenly multiplied by $\alpha$. So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{\eta}$. Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{\eta})$. Thus, $\mathbf{E}[Y_{n}^\alpha] = \frac{\mathbf{E}[Y_{n}]}{\alpha}$, i.e., the average total time required to find $n$ blocks will be divided by $\alpha$, while $\mathbf{V}[Y_{n}^\alpha] = \frac{\mathbf{V}[Y_n]}{\alpha^2}$ and the variance will be divided by $\alpha^2$. Hence, on one hand, when the network's hash rate increases ($\alpha > 1$), the 2016 blocks will be found earlier. On the other hand, when the network's hash rate decreases ($\alpha < 1$), the 2016 blocks will be found later.

@@

@@

It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a$, which yields $\int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
$$F_T(t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$
$$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$

@@

Let $\mathcal{L}$ be the Laplace Transform. Thus,
\begin{align*}
\mathcal{L}\{F_T(t)\} &= \mathcal{L}\{1 - (1+bt)^n e^{-\lambda at}\} \\
	&= \mathcal{L}\{1\} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \tag{$\mathcal{L}$ is a linear operator} \\
	&= \frac{1}{s} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \mathcal{L}\{t^k e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \frac{k!}{(s+\lambda a)^{k+1}}
\end{align*}

@@

Finally, as $\mathbf{E}[T] = -\mathcal{L}\{f_T(t)\}|_{s=0}$,
$$\mathbf{E}[T] = \frac{1}{\lambda a} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right] \text{, where $n = \frac{\lambda(a-1)}{b}$}$$

@@

In order to compare the hash rate change models, namely (i) suddenly change, (ii) smoothly change, and (iii) linear change, I have applied each of them to the same scenarios. In the first scenario, the hashrate will double in the next week, whereas, in the second scenario, it will halve in the next week.

In both the smoothly change model and the linear change model, I could have calculated the average time between blocks during the whole week. It would not give us much information, because the time between models would be increasing (or reducing) more and more as the days goes by. And we are really interested in the average time between blocks throughout the days, and not the average of the whole week.

@@

I compared both the smoothly change model and the linear change model with the suddenly change model. The difference between them is negligible. Let $\epsilon$ be the maximum absolute error between the models, than $\epsilon < 0.8$ and $\epsilon/H < 0.2\%$, for all intervals. The maximum absolute error between the linear and the suddenly change models can be seen in Figure \ref{fig:bitcoin-H-change-error}.

@@

Therefore, we may conclude that it is reasonable to approximate the average time between blocks using only the suddenly change model in each interval of one hour.

The average time between blocks throughout the days can be seen in Figure \ref{fig:bitcoin-H-change}. It was calculated using the suddenly change model with the hash rate changing linearly during the week.

@@

There are many possible ways to attack the Bitcoin \citep{karame2012two, heilman2015eclipse, bahack2013theoretical, bonneau2016buy, neudecker2015simulation}. In this section, we are interested in a particular attack: the double spending attack.

@@

In the Bitcoin's network, the average number of transactions per block is $s=2,250$, so, the system is stable when $\lambda_{\text{TX}} < 2,250/600 = 3.75$ tx/s. Therefore, $3.75$ is the maximum number of new transactions per second that the Bitcoin's network may handle. When $\lambda_{\text{TX}} > 3.75$ tx/s, the unconfirmed transaction pool starts to grow indefinitely.

@@

$m \ll s$ yields $z_s \rightarrow 1 + 1/m$. Thus, the average number of unconfirmed transactions $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$. In the Bitcoin's network, $m \ll s$ is reached when $\lambda_{\text{TX}} \ll 3.75$ tx/s. In other words, when the number of new transactions per second is way smaller than 3.75 tx/s, the average waiting time of a transaction to be confirmed is 600 seconds, which means, on average, all transactions will be confirmed in the next block.

@@

Therefore, we conclude that the Bitcoin's network capacity is $\lambda_{\text{blocks}} = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.

>>>>

The primary objective of this chapter is to increase the understanding of Bitcoin through mathematical tools.

@@

In the Bitcoin protocol, the given number $A$ is adjusted so that the network would find a new a block every 10 minutes, on average. Suppose that the Bitcoin network is able to calculate $H$ hashes per second --- $H$ is the total hash rate of the network. The time required to find a solution would be $T=X/H$, and $\mathbf{E}(T) = \mathbf{E}(X)/H$ would be the average number of seconds to find a new block. So, the rule of finding a new block every 10 minutes ($\eta = 600$ seconds) --- on average --- leads to the following equation: $\mathbf{E}(T) = \eta = 600$. So, $\mathbf{E}(T) = \mathbf{E}(X)/H = \frac{1}{pH} = \eta = 600 \Rightarrow p = \frac{1}{\eta H}$. Finally, $\mathbf{E}(X) = \eta H$, $\mathbf{E}(T) = \eta$, $\mathbf{V}(X) = (\eta H)^2 - \eta H$, and $\mathbf{V}(T) = \eta^2 - \eta/H$.

The cumulative distribution function (CDF) of $T$ is $\mathbf{P}(T \leq t) = \mathbf{P}(X/H \leq t) = \mathbf{P}(X \leq tH) = 1 - (1-p)^{tH} = 1 - \left( 1-\frac{1}{\eta H} \right)^{tH}$. But, as the Bitcoin network hash rate is really large, we may approximate the CDF of $T$ by $\lim_{H \rightarrow \infty} \mathbf{P}(T \leq t) = 1 - e^{-\frac{t}{\eta}}$, which is equal to the CDF of the exponential distribution with parameter $\lambda = \frac{1}{\eta}$.

@@

Bitcoin peer-to-peer network is a gossip network, where miners are semi-randomly connected to each other, and each miner sends all information it receives to all its peers. According to \citet{decker2013information}, the average time for a new block propagate over the network is 12.6 seconds, while the 95\% percentile is 40 seconds, which indicates a long-tail distribution. \citet{bitcoinstats} has measured the propagation time between 2013 and 2017. During 2017, the worst daily 90\% percentile was 21 seconds. Notice that both results may not be contradictory because the Bitcoin network is continuously evolving.

@@

Bitcoin network is open and miners can come and go as they wish, and their movements affect the network's total hash rate. From the theorems developed so far, we can predict how the network behaves in a constant hash rate scenario. But how a change in hash rate affects the network is still to be explored, and that is the goal of this section.

@@

The most simple model for a change in the network's total hash rate is to consider a sudden shift as if it has jumped to a new value without a continuous transition to it.

Let's say that the network's total hash rate has been suddenly multiplied by $\alpha$. So, $u(t) = \alpha$, $\int_0^t u(t) dt = \alpha t$, and $T$ also follows an exponential distribution, but with $\lambda = \frac{\alpha}{\eta}$. Thus, $Y_n^\alpha = \sum_{i=1}^{n} T_i^\alpha \sim \text{Erlang}(n, \frac{\alpha}{\eta})$. Thus, $\mathbf{E}[Y_{n}^\alpha] = \frac{\mathbf{E}[Y_{n}]}{\alpha}$, i.e., the average total time required to find $n$ blocks will be divided by $\alpha$, while $\mathbf{V}[Y_{n}^\alpha] = \frac{\mathbf{V}[Y_n]}{\alpha^2}$ and the variance will be divided by $\alpha^2$. Hence, on one hand, when the network's hash rate increases ($\alpha > 1$), the 2016 blocks will be found earlier. On the other hand, when the network's hash rate decreases ($\alpha < 1$), the 2016 blocks will be found later.

@@

Now let's model a change in the network's total hash rate using a continuous and smoothly function. In this model, the hash rate is multiplied by a function that continuously goes from $1$ towards a new value $a$ over time. The smoothness of the transition --- or how fast it changes --- is also parameterized.

@@

It is easy to integrate $u(t)$ because $\frac{1+abx}{1+bx} = \frac{1-a}{1+bx} + a$, which yields $\int_0^t u(x) dx = at + \frac{1-a}{b} \log(1+bt)$. So,
$$F_T(t) = \mathbf{P}(T \leq t) = 1 - (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$

Let $f_T(t) = \frac{d}{dt} F_T(t)$ be the probability density function, then:

$$f_T(t) = \lambda \left( \frac{1+abt}{1+bt} \right) (1+bt)^{\frac{\lambda(a-1)}{b}} e^{-\lambda at}.$$

@@

To understand how the change in the hash rate affects the time between blocks, we will calculate the average time between blocks, i.e., $\mathbf{E}[T]$. For this, let's use the well-known result that the expected value of a random variable is equal to minus the derivative of the Laplace transformation of its probability density function when $s=0$.

Let $\mathcal{L}$ be the Laplace Transform operator, then $\mathbf{E}[T] = - \frac{d}{ds} \mathcal{L}\{f_T(t)\}|_{s=0}$. First, we will calculate $\mathcal{L}\{F_T(t)\}$, and then we obtain $\mathcal{L}\{f_T(t)\}$ using $\mathcal{L}\{F_T(t)\}$. Thus,

\begin{align*}
\mathcal{L}\{F_T(t)\} &= \mathcal{L}\{1 - (1+bt)^n e^{-\lambda at}\} \\
	&= \mathcal{L}\{1\} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \tag{$\mathcal{L}$ is a linear operator} \\
	&= \frac{1}{s} - \mathcal{L}\{(1+bt)^n e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \mathcal{L}\{t^k e^{-\lambda at}\} \\
	&= \frac{1}{s} - \sum_{k=0}^n \binom{n}{k} b^k \frac{k!}{(s+\lambda a)^{k+1}}
\end{align*}

@@

Finally, as $\mathbf{E}[T] = - \frac{d}{ds} \mathcal{L}\{f_T(t)\}|_{s=0}$,
$$\mathbf{E}[T] = \frac{1}{\lambda a} \left[ n! \left(\frac{b}{\lambda a}\right)^n \sum_{k=0}^n \frac{1}{k!} \left( \frac{\lambda a}{b} \right)^{k} \right] \text{, where $n = \frac{\lambda(a-1)}{b}$}$$

@@

In order to compare the hash rate change models, namely (i) suddenly changing, (ii) smoothly changing, and (iii) linearly changing, I have applied each of them to the same scenarios. In the first scenario, the hashrate will double in the next week, whereas, in the second scenario, it will halve in the next week.

In both the smoothly change model and the linear change model, I could have calculated each model's average time between blocks during one week. But, it would not give us much information, because the estimated average time between models would be increasing (or decreasing) more and more as the days goes by. And we are really interested in the average time between blocks throughout the days, and not the average of one week.

@@

I compared both the smoothly change model and the linear change model with the suddenly changing model. The difference between them is negligible. Let $\epsilon$ be the maximum absolute error between the models, than $\epsilon < 0.8$ and $\epsilon/H < 0.2\%$, for all intervals. The maximum absolute error between the linear and the suddenly changing models can be seen in Figure \ref{fig:bitcoin-H-change-error}.

@@

Therefore, we may conclude that it is reasonable to approximate the average time between blocks using only the suddenly changing model in each interval of one hour.

The average time between blocks throughout the days can be seen in Figure \ref{fig:bitcoin-H-change}. It was calculated using the suddenly changing model with the hash rate changing linearly during the week.

@@

There are many possible ways to attack the Bitcoin network \citep{karame2012two, heilman2015eclipse, bahack2013theoretical, bonneau2016buy, neudecker2015simulation}. In this section, we are interested in a particular attack: the double spending attack.

@@

In the Bitcoin network, the average number of transactions per block is $s=2,250$, so, the system is stable when $\lambda_{\text{TX}} < 2,250/600 = 3.75$ tx/s. Therefore, $3.75$ is the maximum number of new transactions per second that the Bitcoin network may handle. When $\lambda_{\text{TX}} > 3.75$ tx/s, the unconfirmed transaction pool starts to grow indefinitely.

@@

$m \ll s$ yields $z_s \rightarrow 1 + 1/m$. Thus, the average number of unconfirmed transactions $\mathbf{E}(\pi) \rightarrow m$ and the average waiting time $\mathbf{E}(w) \rightarrow \frac{1}{\lambda_{\text{blocks}}} = \eta = 600 \text{ seconds}$. In the Bitcoin network, $m \ll s$ is reached when $\lambda_{\text{TX}} \ll 3.75$ tx/s. In other words, when the number of new transactions per second is way smaller than 3.75 tx/s, the average waiting time of a transaction to be confirmed is 600 seconds, which means, on average, all transactions will be confirmed in the next block.

@@

Therefore, we conclude that the Bitcoin network capacity is $\lambda_{\text{blocks}} = s/\eta = s/600$ transactions per second, where $s$ is the average number of transactions per block.
